---
title: Scrape a Website I
author: Grant Esparza
date: October, 26 2018
output: pdf_document
---

```{r, echo=FALSE, message=FALSE}

library(stringr)
library(httr)
library(dplyr)
library(gridExtra)
library(knitr)
library(pander)
library(data.table)

```

# Introduction

Several of our last assignments have used data from the EPA to explore the
concept of webscraping. While plotting mean Air Quality Index values across
counties, I noticed that Montana had especially high values in western counties
during the year 2016. After learning that Montan had gone through major
wildfires that year, I thought it would be useful to explore the history
of the relationship between AQI and wildfires in this area. I'll be focusing on
data from the years 2005-2015.

****

# Preparing the data

The goal here is to aggreate information about AQI and wildfires so that I have
one complete dataset I can use to do some analysis. I've gone through the steps
of collecting each dataset individually so now I need to put it all together.

## Wildfires

For the data on wildfires, I'll need to connect to my database and extract
information about Montana:

```{r}

## Extract fire table from database
db <-dbConnect(SQLite(), "fire_database.sqlite")
res <- dbSendQuery(db, "SELECT * FROM Fires WHERE State == 'MT'")
fires <- dbFetch(res)

```

Next I'll need to generate the statistics that I'm interested in. For this
analysis I'll need the total number of fires and the mean fire size. I'll
also need to make sure I group by the year and each county.

```{r}

## Get a tibble
montana.fires <- fires %>%
    group_by(FIRE_YEAR, FIPS_NAME) %>%
    summarise(num_fires=n(), mn_fire_size=mean(FIRE_SIZE),
              sum_fire_size=sum(FIRE_SIZE))

```

## AQI

First I need to grab the data from the EPA website. I want data from several years
so I need to save these `csv` files into a folder.

```{r, eval=FALSE}

## Incomplete web links
get_request <- GET("https://aqs.epa.gov/aqsweb/airdata/download_files.html#AQI")
html <- content(get_request, "text")
incomplete_links <- as.data.frame(
                      str_match_all(
                        html, pattern="<a href=\"daily.*county_[0-9]{4}\\.zip\">"
                      ) ## str_match_all
                    ) ## as.data.frame

## Split
incomp_links <- as.data.frame(str_split_fixed(incomplete_links[,1],
                                              pattern="\"", 3))
colnames(incomp_links) <- c("href", "link", "end")

## Add home url
complete_links <- paste("https://aqs.epa.gov/aqsweb/airdata/",
                        incomp_links$link, sep="")

## Add data to folder
for (link in complete_links[4:14]) {
  get_request <- GET(link)
  bin_data <- content(get_request, "raw")
  writeBin(bin_data, "daily-county-aqi")
  unzip(zipfile="daily-county-aqi", exdir="aqi")
}

```

Next I need to combine these files into one dataframe. I'll make use of the
`data.table` function `fread()` to speed up the process.

```{r}

## Read csvs
files <- list.files(path="aqi", full.names=TRUE)

## Stack datasets
aqi.df <- rbindlist(lapply(files, fread))

## Tidy the data
colnames(aqi.df) <- c("region", "subregion", "state.code", "county.code", "date",
                      "aqi", "category", "defining.parameter", "defining.site",
                      "number.of.sites.reporting")
aqi.df$date <- year(aqi.df$date)

```

Now that I have the complete table, I'll generate some statistics.

```{r}

## Get a tibble
montana.aqi <- aqi.df %>%
  filter(region == "Montana") %>%
  group_by(date, subregion) %>%
  summarise(mn_aqi=mean(aqi), md_aqi=median(aqi))

```

## Bringing it all together

Now I'll join the datasets so I have one complete source of data. I'll need to
make sure that they have matching column names so that they join properly.

```{r}

## Rename columns for join
colnames(montana.fires)[colnames(montana.fires) == "FIRE_YEAR"] <- "date"
colnames(montana.fires)[colnames(montana.fires) == "FIPS_NAME"] <- "subregion"

## Join
montana.comp <- montana.aqi %>%
  inner_join(montana.fires, by=c("subregion", "date"))

montana.comp

```

****

# Analysis

Now that everything is in place I'd like to compare number of fires per year in
each county to the mean AQI each year per county. Remember the orignal focus of
this exploration was the western counties of Montana, so I'll need to make sure
my analysis considers only these counties.

## Plot the data

```{r, message=FALSE, warning=FALSE}

## Select western counties
montana.western <- montana.comp %>%
  filter(subregion %in% c("Lincoln", "Flathead", "Sanders", "Lake", "Missoula",
                        "Ravalli", "Powell"))

## Number of fires explains mean_aqi
p <- ggplot(montana.western, aes(log(num_fires), log(mn_aqi))) +
  geom_point(aes(shape=subregion)) +
  stat_smooth(method="lm", se=FALSE) +
  ggtitle("Fires vs AQI per year in Montana counties") +
  xlab("Log of fires per year") + ylab("Log of mean AQI per year") +
  theme_minimal() +
  theme(axis.title.x = element_text(face = "italic"),
        axis.title.y = element_text(face = "italic"),
        axis.text.x = element_text(angle=45, hjust=1),
        title = element_text(face = "bold"))

## Add margin plots
ggMarginal(p, type="boxplot")

```

The plot seems to show a postive relationship between the number of fires and
the mean AQI per year in each county.

## Model

Now that I have reason to believe there is a relationship, I'll build a linear
model to see how it fits.

```{r}

summary(lm(log(mn_aqi)~log(num_fires), data=montana.western))

```

The adjusted $R^{2}$ of **0.59** tells us that 59% of the variance in the log of the mean AQI per year in each county is explained by the log number of fires each year per county.

For each one percent increase in the number of fires per year in each county, there is a0.43 percent increase in the mean AQI per year in each county.

When the log number of fires per year in each county is equal to zero, the log of the
mean AQI per year in each county is 1.59.

****

# Takeaway

This analysis supports my initial suspicion that the high AQI values in western
Montana were related to large wildfires in the area. The idea that AQI is affected
by wildfires is not a revolutionary concept, but it was rewarding to see my assumptions
backed up by data from multiple sources.
